\chapter{Full Cognitive Dimension Analysis}
\label{chap:cogdim}

\newcommand{\sideways}[1]{
  \parbox[t]{2mm}{{\rotatebox[origin=r]{90}{#1}}}}
\newlength{\fstcollen}
\newlength{\sndcollen}
\setlength{\fstcollen}{0.5cm}
\setlength{\sndcollen}{(\textwidth-\fstcollen-2cm)/4}
\begin{longtable}{r | p{\sndcollen} p{\sndcollen} p{\sndcollen} p{\sndcollen}}
\hline\noalign{\medskip}

\sideways{CD}
 &
Accelerate
 &
Non-unifying
 &
Data families
 &
Type families (only stencil data type)

\\\noalign{\medskip}
\hline\noalign{\medskip}

\sideways{Repetition viscosity}
 &
\textbf{Worst}

Clearly here we have a very high viscosity: each function must be re-written in terms of new syntax and run in different ways.
 &
We have improved the viscosity significantly. The user must only
implement their stencils in one language but they must still change all
the imports and correct type errors.
 &
Data families worsen the viscosity over the import method as we must now
change all the data constructors as opposed to the imports. In real code
there will be more of these than import locations.
 &
\textbf{Best}

Here we have the least repetition viscosity of all the approaches. We
now only need to change the quasi quoter to change the whole
implementation.

\\\noalign{\medskip}

\sideways{Imposed lookahead}
 &
\textbf{Worst}

The user must know ahead of time that they will be writing in two
languages to be sure to minimize duplication of code and structure their
program correctly.
 &
\textbf{Best}

There is practially no imposed lookahead as we can simply swap out the
implementation by importing from different places.
 &
\textbf{Best}

We do not have imposed lookahead as we can easily swap the constructors.
 &
\textbf{Best}

There is little imposed lookahead in theory, though some operations are
currently not supported in the GPU implementation. (TODO: link to more)
Some types may not be supported easily in both implementations so this
should be considered too.

\\\noalign{\medskip}

\sideways{Consistency}
 &
\textbf{Worst}

The syntaxes are different and so are fairly inconsistent. There are,
however, some similarities between the two in their stencil
representation.
 &
Consistency is improved as the syntax is now uniform but types are not
uniform.
 &
The syntax and usage is the same except for changing the constructors
which is inconsistent.
 &
\textbf{Best}

We have eliminated the inconsistency in usage of the data families. Now
the approach is almost entirely consistent except for the types.

\\\noalign{\medskip}

\sideways{Terseness}
 &
\textbf{Worst}

A lot of code is written by the user to cope with the two different
implementations.
 &
Changing requires a fair bit of code to be changed as we may be
importing many different things from the Ypnos libraries and all these
things must be changed.
 &
We require a lot of code to express the swap from CPU to GPU.
 &
\textbf{Best}

The syntax for switching is minimally terse.

\\\noalign{\medskip}

\sideways{Hidden dependencies}
 &
\textbf{Best}

No hidden dependencies. It is very clear and explicit what is going on.
 &
\textbf{Worst}

Many hidden dependencies are introduced as the types of the different
imported functions do not necessarily match. This can cause failures in
many different places on changing the import.
 &
\textbf{Best}

Here the dependencies introduced in the type system by the import
approach have been made explicit by data constructors.
 &
\textbf{Worst}

More hidden dependencies are introduced. Types change without the users knowing
due to different type and constraint families. This might affect some programs.

\\\noalign{\medskip}

\sideways{Abstraction gradient}
 &
\textbf{Best}

The abstraction is at its most basic level -- choosing between Ypnos or
Accelerate. The user must get to grips with these abstractions as a
minimum.
 &
The abstraction level is fairly low as it uses only simple Haskell
constructs.
 &
The user must now be familiar with the idea of an associated data family
and GADT which are quite advanced Haskell type features.
 &

The abstraction here is perhaps highest of all as it uses the most
advanced type features.

\\\noalign{\medskip}

\sideways{Closeness of mapping}
 &
\textbf{Best}

This way we preserve the comonadic nature of the operations in the type
so that it is obvious to the user what is going on.
 &
\textbf{Worst}

The comonadicity is lost due to having to change the types to suit the
accelerate implementation.
 &
\textbf{Worst}

The comonadicity is lost also.
 &
\textbf{Worst}

The comonadicity is lost.

\\\noalign{\medskip}
\hline
\noalign{\medskip}
\caption{Comparison of the different API using cognitive dimensions. \label{tbl:cogcomp}}
\end{longtable}

\chapter{Original Proposal}
\label{chap:prop}

\input{propbody.tex}