\chapter{Proforma}

{\large
\begin{tabular}{p{0.3\textwidth} p{0.7\textwidth}}
  Name:               & \bf Samuel Pattuzzi                                 \\
  College:            & \bf Robinson College                                \\
  Project Title:      & \bf \thetitle \\
  Examination:        & \bf Part II Project, June 2013                      \\
  Word Count:         & \bf \wordcount                                      \\
  Project Originator: & D.A. Orchard                                        \\
  Supervisor:         & D.A. Orchard                                        \\
\end{tabular}
}

\section*{Original Aims of the Project}

Creating a backend to the Ypnos programming language to support GPU accelerated
computation. This will require the implementation of translation from the Ypnos
stencil syntax to GPU code. Furthermore, it will require, as a minimum, two
language primitives to be implemented: \emph{run} and \emph{reduce}. Both the
translation and primitives must be correct and, furthermore, outperform the CPU
implementation.

\section*{Work Completed}

The criteria set out in the proposal were met. A backend consisting of a
compilation stage and two primitives was implemented. The correctness of the
implementation was checked by unit testing and benchmarks showed that the
performance of the GPU backend exceeded that of the CPU. In addition, a
usability analysis of the library was performed.

\section*{Special Difficulties}

None.

\newpage
\section*{Declaration}

I, Samuel Pattuzzi of Robinson College, being a candidate for Part II of the
Computer Science Tripos, hereby declare that this dissertation and the work
described in it are my own work, unaided except as may be specified below, and
that the dissertation does not contain material that has already been used to
any substantial extent for a comparable purpose.

I give permission for my dissertation to be made available in the archive
area of the Laboratory's website.

\begin{tabbing}
\bigskip \\
Signed   \= \\
\medskip \\
Date \> \thedate \\
\end{tabbing}

\cleardoublepage
{
 \hypersetup{linkcolor=black}
\setcounter{tocdepth}{2}
\tableofcontents
}
\cleardoublepage
